Designing Machine Learning Pipeline Toolkit for AutoML Surrogate Modeling Optimization

### 2023-05-25 paper/paper.tex
significance testing  discussion
Signed-off-by: ppalmes <ppalmes@gmail.com>

Table \ref{tab:openmldata} summarizes the statistical features of the 12 OpenML
datasets~\cite{openml} used in the experiment.  Also included are the best
results in a 4-hour time budget  discussed in the review paper
of \cite{zoller2019}.  The comparison involves 5 AutoML approaches, namely:
TPOT~\cite{tpot2016}, Auto-Sklearn~\cite{Feurer:AML2019Chp6,Feurer:NIPS2015},
ATM~\cite{atm2017}, Hyperopt-Sklearn~\cite{bergstra2011nips,bergstra2012}, and
Random Search~\cite{anderson1953}.  Among these approaches, TPOT and ATM have
dominated the rankings in terms of best performance based on the
CV classification errors on the 12 datasets as shown in the last two columns.

@@

Table \ref{tab:singleblock} summarizes the performance of \emph{all-all}
strategy. The metric of performance is based on the average classification
error (\emph{AvgErr}) using 10-fold CV. The ranking is based on its
\emph{AvgErr} performance in comparison to the 5 AutoML approaches~\cite{zoller2019} summarized in Table~\ref{tab:openmldata}.
The \emph{all-all} median overall rank is 1.5 using one-block pipeline.
Its median runtime to
perform one-block exhaustive search is 0.23 hour or 13.8 minutes per dataset.
This result is encouraging because the median runtime is significantly less
compared to the 5 AutoML algorithms with a 4-hour budget.  Closely examining 
the runtime of each dataset, \emph{plants-shape} dataset
requires 4.29 hours to finish while
the rest require at most 1.87 hours to run. Take note, however, that the
comparison for runtime is not a fair comparison, as we did not have the time to repeat the experiments done in ~\cite{zoller2019}
using the same machine with our two-stage strategies. Our experiments were conducted using 2017 Model of MacBook Pro with 2.8 GHz Quad-Core Intel Core i7 and 16 GB of RAM.

@@

Comparing Tables \ref{tab:singleblock}, \ref{tab:oneblock-skrf} and \ref{tab:all-one-single}, 
the \emph{all-one} strategy median rank of 1.5 is
similar to \emph{all-all} and significantly better than
the 2.5 median of \emph{one-all}. The \emph{all-one}
median validation error of 27.75
$\pm$ 3.69 is also slightly better than \emph{all-all} (27.92$\pm$ 2.54).
While both have similar validation errors, the biggest advantage of
\emph{all-one} is its speed which is significantly faster than \emph{all-all}.
Its median time duration for all 12 datasets is just 0.05
hour (3 minutes) compared to the runtime median of 0.23 hour (13.8 minutes)
for \emph{all-all}. Furthermore, the dataset with worst runtime is the
\emph{analcadata-dmf} which requires 0.35 hour or 21 minutes.  This worst
runtime is still significantly less than the 4-hour budget allotted to the 5
AutoML algorithms. It is interesting to note that most solutions in
\emph{all-one} do not employ any feature extraction but only scaling. We
can consider the solutions in \emph{all-one} to be sparsed relative to
the dominant presence of \texttt{noop} in the feature extraction part
of the one-block MLPL.

@@

Inspired by these promising results, the next experiment applies the
\emph{all-one} strategy to two-block MLPL with the results summarized in Table
\ref{tab:twoblocks-all-one}. The \emph{all-one} strategy achieves a 1.0 median
rank which suggests that its performance is at least as good or better than the
best results of the 5 AutoML algorithms with respect to the 12 datasets.
While the runtime median of \emph{all-one} is
2.62 hours, 3 datasets require more than 4 hours to finish.  
In the future, it will be interesting to incorporate the time budget into the 
\emph{all-one} strategy and run the other AutoMLs in the same machine to have fair comparison.

@@

We use the \emph{all-one} strategy as the baseline and extended it with 
PRP and LR surrogates to find out their effect in prediction error
and computation time. While there is a 5\% increase in error by
using PRP-surrogate, the median computation time went down from 2.28
hours to just 4.2 minutes. In spite of the 5\% drop in accuracy,
PRP-surrogate median rank of 1.0 still indicates superior performance
relative to other AutoML approaches.

@@

Among the datasets, the two-stage strategies have consistent low performance in
both \emph{sick} and \emph{soybean} problems.  Both datasets are characterized
by large number of missing values compared to the rest of the datasets.
There are 6064 \emph{NA}s in \emph{sick} and 2337 \emph{NA}s in
\emph{soybean}. The poor performance of the two-stage strategies
can be attributed to the non-implementation of imputation or
interpolation filter in the \texttt{basepipeline} data
cleaning routine. Future experiments
will examine which imputation or interpolation routines
can be used to improve AMLP solutions.

>>>>

Table \ref{tab:openmldata} summarizes the statistical features of the 12 OpenML
datasets~\cite{openml} used in the experiment.  Also included are the best
results in a 4-hour time budget  discussed in the review paper
of \cite{zoller2019}.  The comparison involves 5 AutoML approaches, namely:
TPOT~\cite{tpot2016}, Auto-Sklearn~\cite{Feurer:AML2019Chp6,Feurer:NIPS2015},
ATM~\cite{atm2017}, Hyperopt-Sklearn~\cite{bergstra2011nips,bergstra2012}, and
Random Search~\cite{anderson1953}.  The best results among the 5 AutoML approaches
from each dataset  will be used as the baseline comparison in the results and discussion 
of the succeeding section of the paper. Among these 5 approaches, TPOT and ATM have
dominated the rankings in terms of best performance based on the
CV classification errors on the 12 datasets as shown in the last two columns.

@@

Table~\ref{tab:singleblock} summarizes the performance of \emph{all-all}
strategy. The rank represents the relative performance of the proposed algo
relative to the baseline performance of the best algorithm in Table~\ref{tab:openmldata}.
The sign of $=$, $<$, and $>$ indicate whether the rank performance is significantly different
from the baseline at $\alpha=0.05$ level of significance by t-test. 
For example, rank $1, =$ indicates that the proposed algo is as good as
the baseline in Table~\ref{tab:openmldata}. To differentiate the rank where the baseline
or the proposed algorithm is significantly inferior or superior, the $<$ and $>$ are used, 
respectively. For example,  among the datasets, the proposed algo in Table~\ref{tab:singleblock}
is significantly superior to the baseline algos in \emph{first-order-theo}, \emph{profb}, and \emph{plants-shape}
but significantly inferior in \emph{sick} and \emph{soybean} datasets. The best algos in the baseline
has similar performance with the \emph{all-all} strategy in the rest of the datasets.

The metric of performance is based on the average classification
error (\emph{AvgErr}) using 10-fold CV. The ranking is based on its
\emph{AvgErr} performance in comparison to the 
5 AutoML approaches~\cite{zoller2019} summarized in Table~\ref{tab:openmldata}.
The \emph{all-all} median overall rank is 1 using one-block pipeline.
Its median runtime to perform one-block exhaustive search is 0.23 hour or 13.8 minutes per dataset.
This result is encouraging because the median runtime is significantly less
compared to the 5 AutoML algorithms with a 4-hour budget.  Closely examining 
the runtime of each dataset, \emph{plants-shape} dataset
requires 4.29 hours to finish while
the rest require at most 1.87 hours to run. Take note, however, that the
comparison for runtime is not a fair comparison, as we did not have the time to repeat the experiments done in ~\cite{zoller2019}
using the same machine with our two-stage strategies. Our experiments were conducted using 2017 Model of MacBook Pro with 2.8 GHz Quad-Core Intel Core i7 and 16 GB of RAM.

@@

Comparing Tables \ref{tab:singleblock}, \ref{tab:oneblock-skrf} and \ref{tab:all-one-single}, 
the \emph{all-one} strategy median rank of 1 is
similar to \emph{all-all} and \emph{one-all} but the \emph{all-one}
median validation error of 27.75 $\pm$ 3.69 is better than \emph{all-all} and \emph{one-all}.
The biggest advantage of \emph{all-one}
is its speed which is significantly faster to both \emph{all-all} and \emph{one-all}.
Its median time duration for all 12 datasets is just 0.05
hour (3 minutes) compared to the runtime median of 0.23 hour (13.8 minutes)
for \emph{all-all} and 0.11 hour (6 minutes) for \emph{one-all}.

@@

Furthermore, the dataset with worst runtime in \emph{all-one} is the
\emph{jm1} which requires 0.45 hour or 27 minutes.  This worst
runtime is still significantly less than the 4-hour budget allotted to the 5
AutoML algorithms. It is interesting to note that most solutions in
\emph{all-one} do not employ any feature extraction but only scaling. We
can consider the solutions in \emph{all-one} to be sparsed relative to
the dominant presence of \texttt{noop} in the feature extraction part
of the one-block MLPL.

Inspired by these promising results, the next experiment applies the
\emph{all-one} strategy to two-block MLPL with the results summarized in Table~\ref{tab:twoblocks-all-one}. 
The \emph{all-one} strategy achieves a 1 median rank similar to one-block \emph{all-one} 
but the two-block pipeline has much lower cross-validation error compared to one-block pipeline (25.37 vs 27.75).
Similar to one-block \emph{all-one}, the two-block is significantly superior to the 
baseline algos in \emph{first-order-theo}, \emph{profb}, and \emph{plants-shape}
but significantly inferior in \emph{sick} and \emph{soybean} datasets. The best algos in the baseline
has similar performance with the \emph{all-one} strategy in the rest of the datasets.
While the runtime median of \emph{all-one} is
2.62 hours, 3 datasets require more than 4 hours to finish.  
In the future, it will be interesting to incorporate the time budget into the 
\emph{all-one} strategy and run the other AutoMLs in the same machine to have fair comparison.

@@

We use the \emph{all-one} strategy as the baseline and extended it with 
PRP and LR surrogates to find out their effect in prediction error
and computation time. While there is a 5\% increase in error by
using PRP-surrogate, the median computation time went down from 2.28
hours to just 4.2 minutes. In spite of the 5\% drop in accuracy,
PRP-surrogate median rank of 1 still indicates superior performance
relative to other AutoML approaches.

@@

Among the datasets, the one- and two-stage strategies have consistent low performance in
both \emph{sick} and \emph{soybean} problems.  Both datasets are characterized
by large number of missing values compared to the rest of the datasets.
There are 6064 \emph{NA}s in \emph{sick} and 2337 \emph{NA}s in
\emph{soybean}. The poor performance of the two-stage strategies
can be attributed to the non-implementation of imputation or
interpolation filter in the \texttt{basepipeline} data
cleaning routine. Future experiments
will examine which imputation or interpolation routines
can be used to improve AMLP solutions.
